Advances in Artificial Intelligence for Data Visualization: Developing Computer Vision Models to Automate Reading of Data Plots, with Application to Predictive Model Diagnostics

### 2022-02-10 Thesis/01-chap1.Rmd
give potential methods in intro

The multiple individuals approach avoids the limit of $m$, while provides visual tests with $\alpha$ much smaller than $0.05$. In fact, $\alpha$ decreases exponentially as $K$ increases. With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$.

@@

- Not aim for understanding the thought process
- Not aim for mimicking the human vision mechanism
- May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
- May be able to use computer vision model to approximate how people evaluate lineups
  - Use human data to train model with human selection as target -> mimic the human behaviour
  - Use simulated lineup to train model with actual data plot as target -> assume the actual data plot is the most different one
  - Use simulated data plot to train model with null or not null as target ->

>>>>

The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$. In fact, the lower bound of $p$-value decreases exponentially as $K$ increases. With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$.

@@

- Not aim for understanding the thought process
- Not aim for mimicking the human vision mechanism
- May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
- May be able to use computer vision model to approximate how people evaluate lineups

computer vision model:

- Use human data to train model with human selection as target -> mimic the human behaviour
- Use simulated lineup to train model with actual data plot as target -> assume the actual data plot is the most different one
- Use simulated data plot to train model with null or not null as target -> limited to the alternatives given to the model, and Type I error can not be controlled
- Use simulated data plot to train model, then let the model evaluate each plot of a lineup -> leads to multiple selections in a lineup, and the model does not compare different plots in a lineup
  - Use a very large $m$ to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the $m$ should be
  - Use multiple models as multiple individuals to evaluate lineup -> Which models should be used? How many models should be used?
- Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -> how to build such a system? 
  - The input is $m$ plots, what is the output? The probability of being the most different one? 
  - Then how many of plots should be selected? Top 5? Or cumulative probability greater than a threshold? 
  - How to select the threshold?

### 2022-02-10 Thesis/_book/thesis.tex
give potential methods in intro

The multiple individuals approach avoids the limit of \(m\), while provides visual tests with \(\alpha\) much smaller than \(0.05\). In fact, \(\alpha\) decreases exponentially as \(K\) increases. With just \(4\) individuals and \(20\) data plots in a lineup, the \(p\)-value could be as small as \(0.0001\).

@@

\begin{itemize}
\tightlist
\item
  Not aim for understanding the thought process
\item
  Not aim for mimicking the human vision mechanism
\item
  May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
\item
  May be able to use computer vision model to approximate how people evaluate lineups

  \begin{itemize}
  \tightlist
  \item
    Use human data to train model with human selection as target -\textgreater{} mimic the human behaviour
  \item
    Use simulated lineup to train model with actual data plot as target -\textgreater{} assume the actual data plot is the most different one
  \item
    Use simulated data plot to train model with null or not null as target -\textgreater{}
  \end{itemize}
\end{itemize}

>>>>

The multiple individuals approach avoids the limit of \(m\), while provides visual tests with \(p\)-value much smaller than \(0.05\). In fact, the lower bound of \(p\)-value decreases exponentially as \(K\) increases. With just \(4\) individuals and \(20\) data plots in a lineup, the \(p\)-value could be as small as \(0.0001\).

@@

\begin{itemize}
\tightlist
\item
  Not aim for understanding the thought process
\item
  Not aim for mimicking the human vision mechanism
\item
  May be able to define distance metrics to measure difference between data plots for making mathematically prefect decisions
\item
  May be able to use computer vision model to approximate how people evaluate lineups
\end{itemize}

computer vision model:

\begin{itemize}
\tightlist
\item
  Use human data to train model with human selection as target -\textgreater{} mimic the human behaviour
\item
  Use simulated lineup to train model with actual data plot as target -\textgreater{} assume the actual data plot is the most different one
\item
  Use simulated data plot to train model with null or not null as target -\textgreater{} limited to the alternatives given to the model, and Type I error can not be controlled
\item
  Use simulated data plot to train model, then let the model evaluate each plot of a lineup -\textgreater{} leads to multiple selections in a lineup, and the model does not compare different plots in a lineup

  \begin{itemize}
  \tightlist
  \item
    Use a very large \(m\) to ensure the null hypothesis can be rejected by a classifier slightly better than a random selector. The problem becomes how large the \(m\) should be
  \item
    Use multiple models as multiple individuals to evaluate lineup -\textgreater{} Which models should be used? How many models should be used?
  \end{itemize}
\item
  Let the model select multiple plots from a simulated lineup while comparing different plots in a lineup -\textgreater{} how to build such a system?

  \begin{itemize}
  \tightlist
  \item
    The input is \(m\) plots, what is the output? The probability of being the most different one?
  \item
    Then how many of plots should be selected? Top 5? Or cumulative probability greater than a threshold?
  \item
    How to select the threshold?
  \end{itemize}
\end{itemize}
